---
title: "Question 4"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(rstanarm)
library(bayesplot)
library(broom.mixed)
library(ggcorrplot)
```

## Introduction

This markdown file examines how behavioral factors such as physical activity, sedentary behavior, and dietary habits relate to obesity prevalence in the United States. The analysis uses Bayesian hierarchical regression models to evaluate how behavioral indicators are associated with obesity prevalence while accounting for variation across states and time.

## Dataset Formation

```{r}
# Load the raw BRFSS obesity behavior dataset
df <- read_csv("data/raw/adult_obesity_behavior.csv", show_col_types = FALSE)
```

```{r}
# Keep columns needed for modeling purposes only
df_use <- df %>%
  filter(Total == "Total") %>%
  select(LocationDesc, YearStart, Question, Data_Value)
```

```{r}
# Convert long-format survey responses into a wide dataset of behavioral indicators
df_wide <- df_use %>%
  pivot_wider(
    names_from = Question,
    values_from = Data_Value
  )
```

```{r, echo=FALSE}
# Replace long survey question text with shorter variable names
df_wide <- df_wide %>%
  rename(
    obesity = `Percent of adults aged 18 years and older who have obesity`,
    
    overweight = `Percent of adults aged 18 years and older who have an overweight classification`,
                  
    activity_150 = `Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)`,
                    
    activity_150_strength = `Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic physical activity (or an equivalent combination) and engage in muscle-strengthening activities on 2 or more days a week`,
                             
    activity_300 = `Percent of adults who achieve more than 300 minutes a week of moderate-intensity aerobic physical activity or 150 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)`,
                    
    muscle_strengthening = `Percent of adults who engage in muscle-strengthening activities on 2 or more days a week`,
                        
    no_leisure_activity = `Percent of adults who engage in no leisure-time physical activity`,
                            
    fruit_lt_1_per_day = `Percent of adults who report consuming fruit less than one time daily`,
                          
    vegetables_lt_1_per_day = `Percent of adults who report consuming vegetables less than one time daily`
  )
```

```{r}
# Inspect structure and missingness of the cleaned dataset
# NOTE: survey question text was replaced with shorter variables
df_wide %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to = "variable",
               values_to = "missing") %>%
  arrange(desc(missing))
```

```{r}
# Remove rows where all behavioral indicators and obesity measures are missing
# Create the cleaned master dataset used for EDA and model-specific subsets
df_final <- df_wide %>%
  filter(
    !if_all(
      c(
        obesity,
        overweight,
        activity_150,
        activity_150_strength,
        activity_300,
        muscle_strengthening,
        no_leisure_activity,
        fruit_lt_1_per_day,
        vegetables_lt_1_per_day
      ),
      is.na
    )
  )
```

## Exploratory Data Analysis

### Distribution of Behavioral and Obesity Indicators

```{r}
# Select obesity and behavioral indicators and reshape data for faceted visualization
df_final %>%
  select(
    obesity,
    overweight,
    no_leisure_activity,
    activity_150,
    muscle_strengthening,
    fruit_lt_1_per_day,
    vegetables_lt_1_per_day
  ) %>%
  pivot_longer(
    everything(),
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Replace variable names with readable labels for plotting
  mutate(
    variable = recode(variable,
      obesity = "Obesity (%)",
      overweight = "Overweight (%)",
      no_leisure_activity = "No Leisure-Time Activity (%)",
      activity_150 = "Meets 150 Min Activity (%)",
      muscle_strengthening = "Muscle Strengthening ≥2 Days (%)",
      fruit_lt_1_per_day = "Fruit < 1 Time / Day (%)",
      vegetables_lt_1_per_day = "Vegetables < 1 Time / Day (%)"
    )
  ) %>%
  # Plot histograms to visualize distributions of obesity and behavioral indicators
  ggplot(aes(x = value)) +
  geom_histogram(bins = 25, na.rm = TRUE) +
  facet_wrap(~variable, scales = "free") +
  labs(
    title = "Distributions of Behavioral and Obesity Indicators",
    x = "Percent of Adults",
    y = "Number of Location-Year Observations"
  )
```

To gauge the variability of behavioral indicators and obesity across geographic regions, we first examine the distributions of each feature. 

Overall, obesity prevalence across location-year observations ranges roughly from 20% to just over 40%, with most observations concentrated in the 30-35% range. In terms of physical factors, the proportion of adults that meet the required 150 minutes of weekly physical activity ranges from roughly 20-70%, whereas participation in muscle-strengthening activities is lower overall and more concentrated at 25-35%. Across all locations, leisure-time inactivity tends to fall between 15-35% across locations. 

There was also meaningful variation across dietary indicators. The percentage of adults consuming fruit or vegetables once per day generally ranged from 30-50% for fruit and 15-25% for vegetables, which implies differing dietary habits across states and years. 

Overall, the distributions suggest that lifestyle behaviors vary noticeably across the dataset, providing useful variation for examining potential relationships with obesity prevalence.

### Relationship Between Physical Inactivity and Obesity

```{r}
ggplot(df_final, aes(no_leisure_activity, obesity)) +
  geom_point(alpha = 0.4, na.rm = TRUE) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Obesity vs No Leisure-Time Physical Activity",
    x = "No Leisure-Time Physical Activity (%)",
    y = "Obesity (%)"
  )
```

To further explore the relationship between behavioral indicators and obesity prevalence, we examine the association between the percentage of adults reporting no leisure-time physical activity and obesity prevalence. 

The scatter plot shows a clear positive relationship between no leisure-time activity and obesity prevalence. Although there is some variability across observations, the fitted linear trend indicates that obesity prevalence generally increases as the proportion of adults who report no leisure-time physical activity increases. This trend suggests that physical inactivity may be an important factor associated with high obesity prevalence across states and years.

### Relationship Between Physical Activity and Obesity

```{r}
ggplot(df_final, aes(activity_150, obesity)) +
  geom_point(alpha = 0.6, na.rm = TRUE) +
  geom_smooth(method = "lm", se = FALSE, na.rm = TRUE) +
  labs(
    title = "Obesity vs Meeting 150 Minutes of Physical Activity",
    x = "Adults Meeting 150 Minutes of Activity (%)",
    y = "Obesity (%)"
  )
```

After, we examine the relationship between obesity prevalence and the percentage of adults who meet the CDC's recommended 150 minutes of physical activity per week. The scatter plot reveals a negative relationship between these variables, suggesting that locations with higher proportions of adults that adhere to the recommended physical activity guideline exhibit lower obesity prevalence. 

Similar to the previous scatter plot, there is some variation across observations, but the fitted trend line indicates that obesity prevalence generally decreases as the proportions of adults meeting the physical activity guideline increases. 

### Relationship Between Fruit Consumption and Obesity

```{r}
ggplot(df_final, aes(fruit_lt_1_per_day, obesity)) +
  geom_point(alpha = 0.6, na.rm = TRUE) +
  geom_smooth(method = "lm", se = FALSE, na.rm = TRUE) +
  labs(
    title = "Obesity vs Fruit Consumption < 1 Time per Day",
    x = "Fruit < 1 Time / Day (%)",
    y = "Obesity (%)"
  )
```

Additionally, we examine the relationship between obesity prevalence and the percentage of adults who report consuming fruit less than one time per day. The scatter plot has a positive relationship between the two variables, suggesting that locations with higher proportions of adults who consumed fruit less than once daily tend to exhibit higher obesity prevalence. 

The scatter plot shows more mild variability, suggesting that obesity prevalence tends to increase as low fruit consumption is more prevalent. Furthermore, this pattern suggests that poorer dietary habits may be associated with higher obesity prevalence across locations and years. 

### Correlation Between Behavioral Indicators

```{r}
# Compute correlation matrix for obesity and behavioral indicators
corr <- df_final %>%
  select(
    obesity,
    overweight,
    no_leisure_activity,
    activity_150,
    muscle_strengthening,
    fruit_lt_1_per_day,
    vegetables_lt_1_per_day
  ) %>%
  cor(use = "pairwise.complete.obs")

# Replace variable names with readable labels for the heatmap
colnames(corr) <- c(
  "Obesity",
  "Overweight",
  "No Leisure Activity",
  "150 Min Activity",
  "Muscle Strengthening",
  "Fruit < 1 / Day",
  "Vegetables < 1 / Day"
)

rownames(corr) <- colnames(corr)

# Visualize correlations between obesity and behavioral indicators
ggcorrplot(
  corr, 
  lab = TRUE, 
  colors = c("#4575b4", "white", "#d73027")
)
```

To summarize relationship among all behavioral indicators simultaneously, we implemented a correlation matrix shown above. Obesity prevalence shows a moderate positive correlation with leisure-time inactivity (0.47). Furthermore, obesity is also strongly positively correlated with the proportion of adults consuming fruit less than once per day (0.71).

Obesity prevalence shows a negative correlation with the percentage of adults meeting the recommended 150 minutes of physical activity per week (-0.31). Additionally, there is a strong negative correlation between meeting the 150-minute activity guideline and leisure-time inactivity (-0.86). 

These relationships reinforce the patterns observed in the earlier scatter plots, and suggest that both physical activity and dietary behaviors are associated with obesity prevalence. 

### Trends in Obesity Prevalence Over Time

```{r}
# Plot the trend in average obesity prevalence across years
df_final %>%
  group_by(YearStart) %>%
  summarise(mean_obesity = mean(obesity, na.rm = TRUE)) %>%
  ggplot(aes(YearStart, mean_obesity)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(min(df$YearStart), max(df$YearStart), by = 2)) +
  labs(
    title = "Average Obesity Prevalence Over Time",
    x = "Year",
    y = "Average Obesity (%)"
  )
```

Lastly, we examine how obesity prevalence has changed in the last decade. The plot shows that average obesity prevalence rose from 27-28% in the early 2010s to roughly 34% in more recent years. 

## Model Building and Conclusion

### Overview of Models

To examine the relationship between behavioral factors and obesity prevalence, a series of Bayesian linear regression models were fit with obesity prevalence (%) as the response. Each observation corresponds to a state-year pair.

Because indicators were not collected during every survey year, separate models were were estimated for physical inactivity, physical activity, and dietary indicators in order to evaluate how each behavioral indicator relates to obesity prevalence.

### Posterior Inference

Posterior distributions of the model parameters were estimated using Markov Chain Monte Carlo (MCMC) sampling via Stan. Convergence diagnostics and posterior predictive checks were used to assess model fit and ensure that the models adequately captured patterns in the observed data.

### Inactivity Model 

The inactivity indicator corresponds to the variable measuring the percentage of adults who report **no leisure-time physical activity**. In this analysis, this variable is referred to as **inactivity** for brevity.

$obesity_{ij} = \beta_0 + \beta_1 inactive_{ij} + \beta_2 year_j + b_{state[i]} + \epsilon_{ij}$

Where $i$ = state and $j$ = year

$\beta_0$: Intercept representing the expected obesity prevalence when predictors are at their mean values

$\beta_1$: Change in obesity prevalence associated with a one standard deviation increase in the rate of no leisure-time physical activity

$\beta_2$: Association between survey year and obesity prevalence, capturing overall temporal trends

$b_{state[i]}$: State-level random intercept capturing unobserved differences in baseline obesity prevalence across states

```{r}
# Prepare data for inactivity model and standardize predictors
df_inactive <- df_final %>%
  select(LocationDesc, YearStart, obesity, no_leisure_activity) %>%
  drop_na() %>%
  mutate(
    year_z = as.numeric(scale(YearStart)),
    inactive_z = as.numeric(scale(no_leisure_activity))
  )
```

```{r}
# Fit hierarchical Bayesian regression with state-level random intercepts
fit_inactive <- stan_glmer(
  obesity ~ inactive_z + year_z + (1 | LocationDesc),
  data = df_inactive,
  family = gaussian(),
  prior_intercept = normal(30, 10, autoscale = TRUE),
  prior = normal(0, 2, autoscale = TRUE),
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 4,
  iter = 4000,
  seed = 84375,
  refresh = 0
)
```

```{r}
# Extract fixed-effect posterior estimates and 95% credible intervals
broom.mixed::tidy(fit_inactive, effects = "fixed", conf.int = TRUE)
```

The posterior mean for the no leisure-time physical inactivity coefficient is approximately 0.24 with a 95% credible interval of roughly [0.07, 0.41], which does not include 0. This suggests evidence of a positive association between the proportion of adults reporting no leisure-time physical activity and obesity prevalence. 

Additionally, the coefficient for year is also positive with a posterior mean of around 2.17 and a 95% credible interval of around [2.10, 2.24], indicating that obesity prevalence tends to increase over time across states. 

```{r}
# Compute posterior probability that inactivity is positively associated with obesity
post <- as.matrix(fit_inactive)
mean(post[, "inactive_z"] > 0)
```

The posterior probability that the no leisure-time physical inactivity coefficient is positive is around 0.99, indicating strong evidence that higher inactivity rates are associated with higher obesity prevalence.

```{r}
# Posterior predictive check comparing observed and simulated obesity distributions
pp_check(fit_inactive, plotfun = "dens_overlay")
```

The simulated density curves closely overlap the observed density, suggesting that the inactivity model captures the overall distribution of obesity prevalence fairly well. 

```{r}
# Generate posterior predictions for -1 SD and +1 SD inactivity levels
new_data <- tibble(
  LocationDesc = "National",
  year_z = 0,
  inactive_z = c(-1, 1)
)

pred <- posterior_epred(fit_inactive, newdata = new_data)

# Compute posterior mean predicted obesity prevalence
apply(pred, 2, mean)
# Compute 95% credible intervals for predicted obesity prevalence
apply(pred, 2, quantile, probs = c(.025, .975))
```

Posterior predictions were generated for inactivity levels one standard deviation above and below the mean, while holding the standardized year variable at 0. The model predicts an average obesity prevalence of approximately 30.31% when inactivity is one standard deviation below the mean and roughly 30.8% when inactivity is one standard deviation above the mean. 

### Activity Model 

$obesity_ij$ ~ $\beta_0 + \beta_1 * activity150_{ij} + \beta_2 * strength_{ij} + \beta_3 * year_j + b_{state[i]} + \epsilon_i$

Where $i$ = state and $j$ = year

$\beta_0$: Intercept representing the expected obesity prevalence when predictors are at their mean values

$\beta_1$: Association between meeting the 150-minute activity guideline and obesity prevalence

$\beta_2$: Association between muscle-strengthening activity participation and obesity prevalence

$\beta_3$: Association between survey year and obesity prevalence, capturing overall temporal trends

$b_{state[i]}$: State-level random intercept capturing unobserved differences in baseline obesity prevalence across states.

```{r}
# Prepare data for activity model and standardize predictors
df_activity <- df_final %>%
  select(LocationDesc, YearStart, obesity, activity_150, muscle_strengthening) %>%
  drop_na() %>%
  mutate(
    year_z = as.numeric(scale(YearStart)),
    activity150_z = as.numeric(scale(activity_150)),
    strength_z = as.numeric(scale(muscle_strengthening))
  )
```

``` {r}
# Fit hierarchical Bayesian regression including aerobic activity and muscle strengthening
fit_activity <- stan_glmer(
  obesity ~ activity150_z + strength_z + year_z + (1 | LocationDesc),
  data = df_activity,
  family = gaussian(),
  prior_intercept = normal(30, 10, autoscale = TRUE),
  prior = normal(0, 2, autoscale = TRUE),
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 4,
  iter = 4000,
  seed = 84375,
  refresh = 0
)
```

```{r}
# Extract posterior estimates and 95% credible intervals for fixed effects
broom.mixed::tidy(fit_activity, effects = "fixed", conf.int = TRUE)
```
The coefficient for meeting the 150-minute physical activity guideline is negative, with a posterior mean of around -0.32 and a 95% credible interval of [-0.59, -0.04], which excludes 0. This suggests higher rates of aerobic activity participation are associated with lower obesity. 

The coefficient for muscle-strengthening activity has a posterior mean of around 0.15 but the 95% credible interval [-0.19, 0.49] contains 0, which indicates that the relationship between muscle-strengthening activities and obesity prevalence is uncertain in this model. 

Similar to the inactivity model, the year coefficient is positive with a mean of around 2.06 and 95% credible interval of [1.86, 2.27], suggesting that obesity prevalence has increased over time across states. 

```{r}
# Compute posterior probabilities that activity predictors reduce obesity prevalence
post_activity <- as.matrix(fit_activity)

mean(post_activity[, "activity150_z"] < 0)
mean(post_activity[, "strength_z"] < 0)
```

The posterior probability that the coefficient for aerobic activity participation is negative is roughly 0.97, indicating strong evidence that higher activity levels are associated with lower obesity prevalence. 

On the other hand, the posterior probability that the muscle-strengthening coefficient is negative is approximately 0.22, suggesting weak evidence that muscle-strengthening participation reduces obesity prevalence in this model.  

```{r}
# Posterior predictive check comparing observed and simulated obesity distributions
pp_check(fit_activity, plotfun = "dens_overlay")
```

The simulated density curves closely overlap the observed density, suggesting that the activity model captures the overall distribution of obesity prevalence fairly well. 

```{r}
# Generate posterior predictions for -1 SD and +1 SD activity levels
new_data_activity <- tibble(
  LocationDesc = "National",
  year_z = 0,
  activity150_z = c(-1, -1,  1,  1),
  strength_z    = c(-1,  1, -1,  1)
)

pred_activity <- posterior_epred(fit_activity, newdata = new_data_activity)

# Compute posterior mean predicted obesity prevalence
apply(pred_activity, 2, mean)
# Compute 95% credible intervals for predicted obesity prevalence
apply(pred_activity, 2, quantile, probs = c(.025, .975))
```

Posterior predictions were generated for combinations of aerobic activity and muscle-strengthening participation one standard deviation above and below, while holding the standardized year variable at 0. 

The model predicts the lowest obesity prevalence to at around 29.35% when aerobic activity levels are high and strength training participation participation is low. In contrast, obesity prevalence was projected to be 30.29% when aerobic activity levels are low and strength training participation is high. 

These results suggest that higher aerobic activity participation is associated with lower obesity prevalence, while the effect of muscle-strengthening activity appears weaker and more uncertain.

### Nutrition Model

$obesity_{ij} = \beta_0 + \beta_1 fruit_{ij} + \beta_2 veg_{ij} + \beta_3 year_j + b_{state[i]} + \epsilon_{ij}$

Where $i$ = state and $j$ = year

$\beta_0$: Intercept representing the expected obesity prevalence when predictors are at their mean values

$\beta_1$: Change in obesity prevalence associated with a one standard deviation increase in the proportion of adults consuming fruit less than one time per day

$\beta_2$: Change in obesity prevalence associated with a one standard deviation increase in the proportion of adults consuming vegetables less than one time per day

$\beta_3$: Association between survey year and obesity prevalence, capturing overall temporal trends

$b_{state[i]}$: State-level random intercept capturing unobserved differences in baseline obesity prevalence across states

```{r}
# Prepare data for nutrition model and standardize dietary predictors
df_nutrition <- df_final %>%
  select(LocationDesc, YearStart, obesity, fruit_lt_1_per_day, vegetables_lt_1_per_day) %>%
  drop_na() %>%
  mutate(
    year_z = as.numeric(scale(YearStart)),
    fruit_z = as.numeric(scale(fruit_lt_1_per_day)),
    veg_z = as.numeric(scale(vegetables_lt_1_per_day))
  )
```

```{r}
# Fit hierarchical Bayesian regression relating dietary indicators to obesity prevalence
fit_nutrition <- stan_glmer(
  obesity ~ fruit_z + veg_z + year_z + (1 | LocationDesc),
  data = df_nutrition,
  family = gaussian(),
  prior_intercept = normal(30, 10, autoscale = TRUE),
  prior = normal(0, 2, autoscale = TRUE),
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 4,
  iter = 4000,
  seed = 84375,
  refresh = 0
)
```

```{r}
# Extract posterior estimates and 95% credible intervals for dietary predictors
broom.mixed::tidy(fit_nutrition, effects = "fixed", conf.int = TRUE)
```

The posterior estimate suggests that lower fruit consumption is positively associated with higher obesity prevalence. The coefficient for fruit consumption < 1/day has a posterior mean of approximately 1.2 and a 95% credible interval of roughly [0.57, 1.87], which does not contain 0. This indicates that lower fruit consumption is associated with a higher obesity prevalence.

The posterior mean of vegetable consumption < 1/day has a posterior mean of 0.12, but its 95% credible interval of approximately [-0.48, 0.73] contains 0, suggesting that the relationship between low vegetable consumption and obesity prevalence is more uncertain.

Similar to the previous two models, the coefficient for year has a posterior mean of approximately 0.71 and a 95% credible interval of [0.47, 0.95], suggesting that obesity prevalence tends to increase over time across states. 

```{r}
# Compute posterior probabilities that poorer dietary indicators are associated with higher obesity prevalence
post_nutrition <- as.matrix(fit_nutrition)

mean(post_nutrition[, "fruit_z"] > 0)
mean(post_nutrition[, "veg_z"] > 0)
```

The posterior probability that the coefficient for fruit consumption < 1/day being positive is roughly 0.999, indicating very strong evidence that higher rates of consuming fruit < 1/day are associated with a higher obesity prevalence. 

On the other hand, the posterior probability that the coefficient for vegetable consumption < 1/day being positive is around 0.63, which provides weak evidence that higher rates of consuming vegetables < 1/day are associated with a higher obesity prevalence. 

```{r}
# Posterior predictive check comparing observed and simulated obesity distributions
pp_check(fit_nutrition, plotfun = "dens_overlay")
```

The simulated density curves closely overlap the observed density, suggesting that the nutrition model captures the overall distribution of obesity prevalence fairly well. 

```{r}
# Generate posterior predictions for combinations of fruit and vegetable consumption levels
new_data_nutrition <- tibble(
  LocationDesc = "National",
  year_z = 0,
  fruit_z = c(-1, -1, 1, 1),
  veg_z = c(-1, 1, -1, 1)
)

pred_nutrition <- posterior_epred(fit_nutrition, newdata = new_data_nutrition)

apply(pred_nutrition, 2, mean)
apply(pred_nutrition, 2, quantile, probs = c(.025, .975))
```

Posterior predictions were generated for combinations of fruit and vegetable consumption one standard deviation above and below the mean, while holding the standardized year variable at 0. 

Obesity prevalence was projected to at 30.44% when both vegetable and fruit consumption were high (lower percentage of adults consuming fruits and vegetables < 1/day). 

When both fruit and vegetable consumption were lower, obesity prevalence was predicted to be roughly 33.1%.

Overall, these results show that poorer dietary habits, especially lower fruit consumption, are associated with a higher obesity prevalence.

### Comparing Behavioral Predictors Across Models

```{r}
# Extract posterior estimates and credible intervals from each model
coef_inactive <- tidy(fit_inactive, effects = "fixed", conf.int = TRUE)
coef_activity <- tidy(fit_activity, effects = "fixed", conf.int = TRUE)
coef_nutrition <- tidy(fit_nutrition, effects = "fixed", conf.int = TRUE)

# Label coefficients by model type
coef_inactive$model <- "Inactivity"
coef_activity$model <- "Activity"
coef_nutrition$model <- "Nutrition"

# Combine results across models
coef_all <- bind_rows(coef_inactive, coef_activity, coef_nutrition)

# Remove intercept and year terms to focus on behavioral predictors
coef_all <- coef_all %>%
  filter(!term %in% c("(Intercept)", "year_z"))

# Replace variable names with more concise labels for plotting
coef_all$term <- recode(coef_all$term,
  activity150_z = "Aerobic activity",
  strength_z = "Muscle strengthening",
  inactive_z = "No leisure activity",
  fruit_z = "Low fruit consumption",
  veg_z = "Low vegetable consumption"
)

# Plot posterior estimates and 95% credible intervals across models
ggplot(coef_all, aes(x = estimate, y = term)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~model, scales = "free_y") +
  labs(
    title = "Posterior Estimates of Behavioral Predictors on Obesity",
    x = "Estimated Effect on Obesity Prevalence (95% Credible Interval)",
    y = NULL
  ) +
  theme_minimal()
```

The figure summarizes posterior estimates and 95% credible intervals for behavioral predictors across, nutrition, activity, and inactivity. The points represent estimated posterior means of each predictor, horizontal lines represent the 95% credible interval, and the dashed vertical lines at 0 represent no association with obesity prevalence. 

Overall, the results show that aerobic activity is negatively associated with obesity prevalence, while lower fruit consumption and no physical activity during leisure time are positively associated with obesity prevalence. In contrast, predictors such as muscle strengthening activity and low vegetable consumption contain 0 in their credible intervals, which suggest weaker or more uncertain associations with obesity prevalence. 

## Conclusion

**What behavioral factors are posteriorly associated with obesity prevalence, including physical activity, sedentary behavior, and nutrition?**

The posterior results suggest that several behavioral factors are associated with obesity prevalence. Higher rates of leisure time inactivity and lower fruit consumption were positively associated with obesity prevalence. On the other hand, higher rates of aerobic activity were negatively associated with obesity prevalence. 

Factors such as muscle strengthening and vegetable consumption were more uncertain, providing limited evidence for a clear association between obesity prevalence. 
