---
title: "Exploratory Data Analysis: Adult Lifestyle and Obesity"
author: "Brandon Wu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(ggcorrplot)
library(dplyr)
```

**Data Set Loading and Initial Checks**

```{r}
df <- read_csv("../data/cleaned/obesity_behavior_activity_only.csv")
glimpse(df)
```
```{r}
summary(df)
```
```{r}
nrow(df)
n_distinct(df$LocationDesc)
n_distinct(df$YearStart)
```

The data set contains state-level behavioral indicators and obesity prevalence from 2011 to 2023. Each observation corresponds to a state-year pair and 
reports the percentage of adults exhibiting specific lifestyle behaviors.

**Distribution of Adult Obesity Prevalence**

```{r}
ggplot(df, aes(x = obesity)) +
  geom_histogram(
    bins = 25,
    fill = "#0099F8",
    color = "white"
  ) +
  labs(
    title = "Distribution of State-Level Adult Obesity Prevalence",
    x = "Adult Obesity Prevalence (%)",
    y = "Number of State-Year Observations",
    caption = "Each observation represents a state in a given year (2011–2023)."
  ) +
  theme_minimal(base_size = 13)
```

The distribution of obesity prevalence across-state year observations is centered at roughly 30%. Most observed obesity rates fall within the 27% to 34% range, with few states exhibiting high or low prevalence. 

The relatively narrow distribution suggests that differences in behavioral indicators may help explain variation in obesity prevalence across states.

**Distribution of Adult Aerobic Activity Levels**

```{r}
ggplot(df, aes(aerobic_activity_150)) +
  geom_histogram(
    bins = 25,
    fill = "#0099F8",
    color = "white"
  ) + 
  labs(
    title = "Distribution of Adults Meeting Recommended Aerobic Activity Levels",
    x = "Adults Achieving >=150 Minutes of Aerobic Activity per Week (%)",
    y = "Number of State-Year Obsevations",
    caption = "The 150-minute threshold reflects the CDC’s recommended weekly aerobic activity level for adults."
  ) + 
  theme_minimal()
```

The distribution of adults meeting the recommended >=150 minutes of weekly aerobic activity is roughly centered around 50%, indicating that in most state-years approximately half of adults meet the CDC activity guideline. A small number of observations fall below 40%, suggesting that some states or years exhibit substantially lower physical activity levels.

**Physical Inactivity and Adult Obesity Scatterplot**

```{r}
ggplot(df, aes(no_leisure_physical_activity, obesity)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "#0072B2") +
  labs(
    title = "Relationship Between Physical Inactivity and Adult Obesity",
    x = "Adults Reporting No Leisure-Time Physical Activity (%)",
    y = "Adult Obesity Prevalence (%)",
    caption = "Each point represents a state-year observation."
  ) +
  theme_minimal(base_size = 13)
```

The scatter plot shows a positive relationship between physical inactivity and obesity prevalence. State-years with a higher percentage of adults reporting no leisure-time physical activity tend to have higher obesity rates as well.

**Correlation Heat Map**

```{r, fig.width=12, fig.height=6}
corr_matrix <- df %>%
  select(
    obesity,
    aerobic_activity_150,
    muscle_strengthening_2plus,
    no_leisure_physical_activity
  ) %>%
  rename(
    "Obesity Prevalence" = obesity,
    "Aerobic Activity >=150 min/week" = aerobic_activity_150,
    "Muscle Strengthening >=2 days/week" = muscle_strengthening_2plus,
    "No Leisure-Time Physical Activity" = no_leisure_physical_activity
  ) %>%
  cor()

ggcorrplot(
  corr_matrix,
  lab = TRUE,
  type = "lower",
  colors = c("#B2182B", "white", "#2166AC")
) +
  labs(
    title = "Correlation Between Behavioral Indicators and Obesity Prevalence",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

The correlation analysis shows that physical inactivity has the strongest positive association with obesity prevalence (r ≈ 0.52), suggesting that states with higher inactivity rates tend to have higher obesity levels. Aerobic activity shows a modest negative correlation with obesity (r ≈ -0.31), while muscle strengthening activities appear largely unrelated to obesity prevalence (r ≈ -0.04). Additionally, aerobic activity and muscle strengthening are strongly positively correlated (r ≈ 0.79), indicating that states with higher aerobic participation also tend to report higher rates of strength training.

```{r}
df_long <- df %>%
  select(
    obesity,
    aerobic_activity_150,
    muscle_strengthening_2plus,
    no_leisure_physical_activity
  ) %>%
  pivot_longer(
    cols = -obesity,
    names_to = "behavior",
    values_to = "value"
  )

ggplot(df_long, aes(value, obesity)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "#0072B2") +
  facet_wrap(~behavior, scales = "free_x") +
  labs(
    title = "Relationships Between Behavioral Indicators and Obesity Prevalence",
    x = "Behavioral Indicator (%)",
    y = "Obesity Prevalence (%)"
  ) +
  theme_minimal()
```

