---
title: "Exploratory Data Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(ggcorrplot)
```

## Dataset Overview 

The data set contains state-level behavioral health indicators related to obesity prevalence in the United States and its territories. Each observation represents a location-year combination and includes measures of obesity prevalence and several lifestyle indicators such as physical activity and dietary behaviors. 

The goal of this EDA is to examine how behavioral factors in the data set relate to obesity prevalence.

```{r}
df <- read_csv("../data/cleaned/adult_behavior_panel.csv", show_col_types = FALSE)
```

```{r}
# Count missing values for each variable and sort from most to least missing
df %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to = "variable",
               values_to = "missing") %>%
  arrange(desc(missing))
```

Missingness is concentrated in activity and nutrition variables because these indicators were not collected every year in the survey. As a result, some state-year observations lack measurements for these behavioral indicators while core variables such as obesity prevalence, location, and year remain complete. 

## Distribution of Behavioral and Obesity Indicators

```{r}
# Select obesity and behavioral indicators and reshape data for faceted visualization
df %>%
  select(
    obesity,
    overweight,
    no_leisure_activity,
    activity_150,
    muscle_strengthening,
    fruit_lt_1_per_day,
    vegetables_lt_1_per_day
  ) %>%
  pivot_longer(
    everything(),
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Replace variable names with readable labels for plotting
  mutate(
    variable = recode(variable,
      obesity = "Obesity (%)",
      overweight = "Overweight (%)",
      no_leisure_activity = "No Leisure-Time Activity (%)",
      activity_150 = "Meets 150 Min Activity (%)",
      muscle_strengthening = "Muscle Strengthening ≥2 Days (%)",
      fruit_lt_1_per_day = "Fruit < 1 Time / Day (%)",
      vegetables_lt_1_per_day = "Vegetables < 1 Time / Day (%)"
    )
  ) %>%
  # Plot histograms to visualize distributions of obesity and behavioral indicators
  ggplot(aes(x = value)) +
  geom_histogram(bins = 25, na.rm = TRUE) +
  facet_wrap(~variable, scales = "free") +
  labs(
    title = "Distributions of Behavioral and Obesity Indicators",
    x = "Percent of Adults",
    y = "Number of Location-Year Observations"
  )
```

To gauge the variability of behavioral indicators and obesity across geographic regions, we first examine the distributions of each feature. 

Overall, obesity prevalence across location-year observations ranges roughly from 20% to just over 40%, with most observations concentrated in the 30-35% range. In terms of physical factors, the proportion of adults that meet the required 150 minutes of weekly physical activity ranges from roughly 20-70%, whereas participation in muscle-strengthening activities is lower overall and more concentrated at 25-35%. Across all locations, leisure-time inactivity tends to fall between 15-35% across locations. 

There was also meaningful variation across dietary indicators. The percentage of adults consuming fruit or vegetables once per day generally ranged from 30-50% for fruit and 15-25% for vegetables, which implies differing dietary habits across states and years. 

Overall, the distributions suggest that lifestyle behaviors vary noticeably across the data set, providing useful variation for examining potential relationships with obesity prevalence. 

## Relationship Between Physical Inactivity and Obesity

```{r}
ggplot(df, aes(no_leisure_activity, obesity)) +
  geom_point(alpha = 0.4, na.rm = TRUE) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Obesity vs No Leisure-Time Physical Activity",
    x = "No Leisure-Time Physical Activity (%)",
    y = "Obesity (%)"
  )
```

To further explore the relationship between behavioral indicators and obesity prevalence, we examine the association between the percentage of adults reporting no leisure-time physical activity and obesity prevalence. 

The scatter plot shows a clear positive relationship between no leisure-time activity and obesity prevalence. Although there is some variability across observations, the fitted linear trend indicates that obesity prevalence generally increases as the proportion of adults who report no leisure-time physical activity increases. This trend suggests that physical inactivity may be an important factor associated with high obesity prevalence across states and years. 

## Relationship Between Physical Activity and Obesity

```{r}
ggplot(df, aes(activity_150, obesity)) +
  geom_point(alpha = 0.6, na.rm = TRUE) +
  geom_smooth(method = "lm", se = FALSE, na.rm = TRUE) +
  labs(
    title = "Obesity vs Meeting 150 Minutes of Physical Activity",
    x = "Adults Meeting 150 Minutes of Activity (%)",
    y = "Obesity (%)"
  )
```

After, we examine the relationship between obesity prevalence and the percentage of adults who meet the CDC's recommended 150 minutes of physical activity per week. The scatter plot reveals a negative relationship between these variables, suggesting that locations with higher proportions of adults that adhere to the recommended physical activity guideline exhibit lower obesity prevalence. 

Similar to the previous scatter plot, there is some variation across observations, but the fitted trend line indicates that obesity prevalence generally decreases as the proportions of adults meeting the physical activity guideline increases. 

## Relationship Between Fruit Consumption and Obesity

```{r}
ggplot(df, aes(fruit_lt_1_per_day, obesity)) +
  geom_point(alpha = 0.6, na.rm = TRUE) +
  geom_smooth(method = "lm", se = FALSE, na.rm = TRUE) +
  labs(
    title = "Obesity vs Fruit Consumption < 1 Time per Day",
    x = "Fruit < 1 Time / Day (%)",
    y = "Obesity (%)"
  )
```

Additionally, we examine the relationship between obesity prevalence and the percentage of adults who report consuming fruit less than one time per day. The scatter plot has a positive relationship between the two variables, suggesting that locations with higher proportions of adults who consumed fruit less than once daily tend to exhibit higher obesity prevalence. 

The scatter plot shows more mild variability, suggesting that obesity prevalence tends to increase as low fruit consumption is more prevalent. Furthermore, this pattern suggests that poorer dietary habits may be associated with higher obesity prevalence across locations and years. 

## Correlation Between Behavioral Indicators

```{r}
# Compute correlation matrix for obesity and behavioral indicators
corr <- df %>%
  select(
    obesity,
    overweight,
    no_leisure_activity,
    activity_150,
    muscle_strengthening,
    fruit_lt_1_per_day,
    vegetables_lt_1_per_day
  ) %>%
  cor(use = "pairwise.complete.obs")

# Replace variable names with readable labels for the heatmap
colnames(corr) <- c(
  "Obesity",
  "Overweight",
  "No Leisure Activity",
  "150 Min Activity",
  "Muscle Strengthening",
  "Fruit < 1 / Day",
  "Vegetables < 1 / Day"
)

rownames(corr) <- colnames(corr)

# Visualize correlations between obesity and behavioral indicators
ggcorrplot(
  corr, 
  lab = TRUE, 
  colors = c("#4575b4", "white", "#d73027")
)
```

To summarize relationship among all behavioral indicators simultaneously, we implemented a correlation matrix shown above. Obesity prevalence shows a moderate positive correlation with leisure-time inactivity (0.47). Furthermore, obesity is also strongly positively correlated with the proportion of adults consuming fruit less than once per day (0.71).

Obesity prevalence shows a negative correlation with the percentage of adults meeting the recommended 150 minutes of physical activity per week (-0.31). Additionally, there is a strong negative correlation between meeting the 150-minute activity guideline and leisure-time inactivity (-0.86). 

These relationships reinforce the patterns observed in the earlier scatter plots, and suggest that both physical activity and dietary behaviors are associated with obesity prevalence. 

## Trends in Obesity Prevalence Over Time

```{r}
# Plot the trend in average obesity prevalence across years
df %>%
  group_by(YearStart) %>%
  summarise(mean_obesity = mean(obesity, na.rm = TRUE)) %>%
  ggplot(aes(YearStart, mean_obesity)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(min(df$YearStart), max(df$YearStart), by = 2)) +
  labs(
    title = "Average Obesity Prevalence Over Time",
    x = "Year",
    y = "Average Obesity (%)"
  )
```

Lastly, we examine how obesity prevalence has changed in the last decade. The plot shows that average obesity prevalence rose from 27-28% in the early 2010s to roughly 34% in more recent years. 