---
title: "Model Building"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(rstanarm)
library(bayesplot)
library(broom.mixed)
```

## Overview of Models

To examine the relationship between behavioral factors and obesity prevalence, a series of Bayesian linear regression models were fit with obesity prevalence (%) as the response. Each observation corresponds to a state-year pair.

Because indicators were not collected during every survey year, separate models were were estimated for physical inactivity, physical activity, and dietary indicators in order to evaluate how each behavioral indicator relates to obesity prevalence. 

### Priors

Weakly informative priors were used to stabilize computations and guide parameters toward realistic ranges, while allowing data to dominate the posterior inference. 

The priors used are:

$\beta_j$ ~ $N(0, 10)$

$\sigma$ ~ $Exp(1)$

### Posterior Inference

Posterior distributions of the model parameters were estimated using Markov Chain Monte Carlo (MCMC) sampling via Stan. Convergence diagnostics and posterior predictive checks were used to assess model fit and ensure that the models adequately captured patterns in the observed data.

## Inactivity Model 

```{r}
df <- read_csv("../data/cleaned/adult_behavior_panel.csv", show_col_types = FALSE)
```

$obesity_{ij} = \beta_0 + \beta_1 inactive_{ij} + \beta_2 year_j + b_{state[i]} + \epsilon_{ij}$

Where $i$ = state and $j$ = year

$\beta_0$: Intercept representing the expected obesity prevalence when predictors are at their mean values

$\beta_1$: Change in obesity prevalence associated with a one standard deviation increase in the rate of no leisure-time physical activity

$\beta_2$: Association between survey year and obesity prevalence, capturing overall temporal trends

$b_{state[i]}$: State-level random intercept capturing unobserved differences in baseline obesity prevalence across states

```{r}
# Prepare data for inactivity model and standardize predictors
df_inactive <- df %>%
  select(LocationDesc, YearStart, obesity, no_leisure_activity) %>%
  drop_na() %>%
  mutate(
    year_z = as.numeric(scale(YearStart)),
    inactive_z = as.numeric(scale(no_leisure_activity))
  )
```

```{r}
# Fit hierarchical Bayesian regression with state-level random intercepts
fit_inactive <- stan_glmer(
  obesity ~ inactive_z + year_z + (1 | LocationDesc),
  data = df_inactive,
  family = gaussian(),
  prior_intercept = normal(30, 10, autoscale = FALSE),
  prior = normal(0, 2, autoscale = FALSE),
  prior_aux = exponential(1, autoscale = FALSE),
  chains = 4,
  iter = 4000,
  seed = 84375,
  refresh = 0
)
```

```{r}
# Extract fixed-effect posterior estimates and 95% credible intervals
broom.mixed::tidy(fit_inactive, effects = "fixed", conf.int = TRUE)
```

The posterior mean for the no leisure-time physical inactivity coefficient is approximately 0.24 with a 95% credible interval of roughly [0.07, 0.42], which does not include 0. This suggests evidence of a positive association between the proportion of adults reporting no leisure-time physical activity and obesity prevalence. 

Additionally, the coefficient for year is also positive with a posterior mean of around 2.17 and a 95% credible interval of around [2.10, 2.24], indicating that obesity prevalence tends to increase over time across states. 

```{r}
# Compute posterior probability that inactivity is positively associated with obesity
post <- as.matrix(fit_inactive)
mean(post[, "inactive_z"] > 0)
```

The posterior probability that the no leisure-time physical inactivity coefficient is positive is around 0.99, indicating strong evidence that higher inactivity rates are associated with higher obesity prevalence.

```{r}
# Posterior predictive check comparing observed and simulated obesity distributions
pp_check(fit_inactive, plotfun = "dens_overlay")
```

The simulated density curves closely overlap the observed density, suggesting that the inactivity model captures the overall distribution of obesity prevalence fairly well. 

```{r}
# Generate posterior predictions for -1 SD and +1 SD inactivity levels
new_data <- tibble(
  LocationDesc = "National",
  year_z = 0,
  inactive_z = c(-1, 1)
)

pred <- posterior_epred(fit_inactive, newdata = new_data)

# Compute posterior mean predicted obesity prevalence
apply(pred, 2, mean)
# Compute 95% credible intervals for predicted obesity prevalence
apply(pred, 2, quantile, probs = c(.025, .975))
```

Posterior predictions were generated for inactivity levels one standard deviation above and below the mean, while holding the standardized year variable at 0. The model predicts an average obesity prevalence of approximately 30.3% when inactivity is one standard deviation below the mean and roughly 30.8% when inactivity is one standard deviation above the mean. 

## Activity Model 

$obesity_ij$ ~ $\beta_0 + \beta_1 * activity150_{ij} + \beta_2 * strength_{ij} + \beta_3 * year_j + b_{state[i]} + \epsilon_i$

Where $i$ = state and $j$ = year

$\beta_0$: Intercept representing the expected obesity prevalence when predictors are at their mean values

$\beta_1$: Association between meeting the 150-minute activity guideline and obesity prevalence

$\beta_2$: Association between muscle-strengthening activity participation and obesity prevalence

$\beta_3$: Association between survey year and obesity prevalence, capturing overall temporal trends

$b_{state[i]}$: State-level random intercept capturing unobserved differences in baseline obesity prevalence across states.

```{r}
# Prepare data for activity model and standardize predictors
df_activity <- df %>%
  select(LocationDesc, YearStart, obesity, activity_150, muscle_strengthening) %>%
  drop_na() %>%
  mutate(
    year_z = as.numeric(scale(YearStart)),
    activity150_z = as.numeric(scale(activity_150)),
    strength_z = as.numeric(scale(muscle_strengthening))
  )
```

``` {r}
# Fit hierarchical Bayesian regression including aerobic activity and muscle strengthening
fit_activity <- stan_glmer(
  obesity ~ activity150_z + strength_z + year_z + (1 | LocationDesc),
  data = df_activity,
  family = gaussian(),
  prior_intercept = normal(30, 10, autoscale = FALSE),
  prior = normal(0, 2, autoscale = FALSE),
  prior_aux = exponential(1, autoscale = FALSE),
  chains = 4,
  iter = 4000,
  seed = 84375,
  refresh = 0
)
```

```{r}
# Extract posterior estimates and 95% credible intervals for fixed effects
broom.mixed::tidy(fit_activity, effects = "fixed", conf.int = TRUE)
```
The coefficient for meeting the 150-minute physical activity guideline is negative, with a posterior mean of around -0.311 and a 95% credible interval of [-0.58, -0.04], which excludes 0. This suggests higher rates of aerobic activity participation are associated with lower obesity. 

The coefficient for muscle-strengthening activity has a posterior mean of 0.16 but the 95% credible interval [-0.16, 0.49] contains 0, which indicates that the relationship between muscle-strengthening activities and obesity prevalence is uncertain in this model. 

Similar to the inactivity model, the year coefficient is positive with a mean of around 2.05 and 95% credible interval of [1.85, 2.26], suggesting that obesity prevalence has increased over time across states. 

```{r}
# Compute posterior probabilities that activity predictors reduce obesity prevalence
post_activity <- as.matrix(fit_activity)

mean(post_activity[, "activity150_z"] < 0)
mean(post_activity[, "strength_z"] < 0)
```

The posterior probability that the coefficient for aerobic activity participation is negative is roughly 0.97, indicating strong evidence that higher activity levels are associated with lower obesity prevalence. 

On the other hand, the posterior probability that the muscle-strengthening coefficient is negative is approximately 0.21, suggesting weak evidence that muscle-strengthening participation reduces obesity prevalence in this model.  

```{r}
# Posterior predictive check comparing observed and simulated obesity distributions
pp_check(fit_activity, plotfun = "dens_overlay")
```

The simulated density curves closely overlap the observed density, suggesting that the activity model captures the overall distribution of obesity prevalence fairly well. 

```{r}
# Generate posterior predictions for -1 SD and +1 SD activity levels
new_data_activity <- tibble(
  LocationDesc = "National",
  year_z = 0,
  activity150_z = c(-1, -1,  1,  1),
  strength_z    = c(-1,  1, -1,  1)
)

pred_activity <- posterior_epred(fit_activity, newdata = new_data_activity)

# Compute posterior mean predicted obesity prevalence
apply(pred_activity, 2, mean)
# Compute 95% credible intervals for predicted obesity prevalence
apply(pred_activity, 2, quantile, probs = c(.025, .975))
```

Posterior predictions were generated for combinations of aerobic activity and muscle-strengthening participation one standard deviation above and below, while holding the standardized year variable at 0. 

The model predicts the lowest obesity prevalence to at around 29.35% when aerobic activity levels are high and strength training participation participation is low. In contrast, obesity prevalence was projected to be 30.3% when aerobic activity levels are low and strength training participation is high. 

These results suggest that higher aerobic activity participation is associated with lower obesity prevalence, while the effect of muscle-strengthening activity appears weaker and more uncertain. 

# Nutrition Model

$obesity_{ij} = \beta_0 + \beta_1 fruit_{ij} + \beta_2 veg_{ij} + \beta_3 year_j + b_{state[i]} + \epsilon_{ij}$

Where $i$ = state and $j$ = year

$\beta_0$: Intercept representing the expected obesity prevalence when predictors are at their mean values

$\beta_1$: Change in obesity prevalence associated with a one standard deviation increase in the proportion of adults consuming fruit less than one time per day

$\beta_2$: Change in obesity prevalence associated with a one standard deviation increase in the proportion of adults consuming vegetables less than one time per day

$\beta_3$: Association between survey year and obesity prevalence, capturing overall temporal trends

$b_{state[i]}$: State-level random intercept capturing unobserved differences in baseline obesity prevalence across states

```{r}
# Prepare data for nutrition model and standardize dietary predictors
df_nutrition <- df %>%
  select(LocationDesc, YearStart, obesity, fruit_lt_1_per_day, vegetables_lt_1_per_day) %>%
  drop_na() %>%
  mutate(
    year_z = as.numeric(scale(YearStart)),
    fruit_z = as.numeric(scale(fruit_lt_1_per_day)),
    veg_z = as.numeric(scale(vegetables_lt_1_per_day))
  )
```

```{r}
# Fit hierarchical Bayesian regression relating dietary indicators to obesity prevalence
fit_nutrition <- stan_glmer(
  obesity ~ fruit_z + veg_z + year_z + (1 | LocationDesc),
  data = df_nutrition,
  family = gaussian(),
  prior_intercept = normal(30, 10, autoscale = FALSE),
  prior = normal(0, 2, autoscale = FALSE),
  prior_aux = exponential(1, autoscale = FALSE),
  chains = 4,
  iter = 4000,
  seed = 84375,
  refresh = 0
)
```

```{r}
# Extract posterior estimates and 95% credible intervals for dietary predictors
broom.mixed::tidy(fit_nutrition, effects = "fixed", conf.int = TRUE)
```

The posterior estimate suggests that lower fruit consumption is positively associated with higher obesity prevalence. The coefficient for fruit consumption < 1/day has a posterior mean of approximately 1.17 and a 95% credible interval of roughly [0.55, 1.78], which does not contain 0. This indicates that lower fruit consumption is associated with a higher obesity prevalence.

The posterior mean of vegetable consumption < 1/day has a posterior mean of 0.15, but its 95% credible interval of approximately [-0.42, 0.75] contains 0, suggesting that the relationship between low vegetable consumption and obesity prevalence is more uncertain.

Similar to the previous two models, the coefficient for year has a posterior mean of 0.72 and a 95% credible interval of [0.49, 0.96], suggesting that obesity prevalence tends to increase over time across states. 

```{r}
# Compute posterior probabilities that poorer dietary indicators are associated with higher obesity prevalence
post_nutrition <- as.matrix(fit_nutrition)

mean(post_nutrition[, "fruit_z"] > 0)
mean(post_nutrition[, "veg_z"] > 0)
```

The posterior probability that the coefficient for fruit consumption < 1/day being posttive is roughly 0.999, indicating very strong evidence that higher rates of consuming fruit < 1/day are associated with a higher obesity prevalence. 

On the other hand, the posterior probability that the coefficient for vegetable consumption < 1/day being positive is 0.66, which provides weak evidence that higher rates of consuming vegetables < 1/day are associated with a higher obesity prevalence. 

```{r}
# Posterior predictive check comparing observed and simulated obesity distributions
pp_check(fit_nutrition, plotfun = "dens_overlay")
```

The simulated density curves closely overlap the observed density, suggesting that the nutrition model captures the overall distribution of obesity prevalence fairly well. 

```{r}
# Generate posterior predictions for combinations of fruit and vegetable consumption levels
new_data_nutrition <- tibble(
  LocationDesc = "National",
  year_z = 0,
  fruit_z = c(-1, -1, 1, 1),
  veg_z = c(-1, 1, -1, 1)
)

pred_nutrition <- posterior_epred(fit_nutrition, newdata = new_data_nutrition)

apply(pred_nutrition, 2, mean)
apply(pred_nutrition, 2, quantile, probs = c(.025, .975))
```

Posterior predictions were generated for combinations of fruit and vegetable consumption one standard deviation above and below the mean, while holding the standardized year variable at 0. 

Obesity prevalence was projected to at 30.44% when both vegetable and fruit consumption were high (lower percentage of adults consuming fruits and vegetables < 1/day). 

When both fruit and vegetable consumption were lower, obesity prevalence was predicted to be roughly 33.07%.

Overall, these results show that poorer dietary habits, especially lower fruit consumption, are associated with a higher obesity prevalence.

## Comparing Behavioral Predictors Across Models

```{r}
# Extract posterior estimates and credible intervals from each model
coef_inactive <- tidy(fit_inactive, effects = "fixed", conf.int = TRUE)
coef_activity <- tidy(fit_activity, effects = "fixed", conf.int = TRUE)
coef_nutrition <- tidy(fit_nutrition, effects = "fixed", conf.int = TRUE)

# Label coefficients by model type
coef_inactive$model <- "Inactivity"
coef_activity$model <- "Activity"
coef_nutrition$model <- "Nutrition"

# Combine results across models
coef_all <- bind_rows(coef_inactive, coef_activity, coef_nutrition)

# Remove intercept and year terms to focus on behavioral predictors
coef_all <- coef_all %>%
  filter(!term %in% c("(Intercept)", "year_z"))

# Replace variable names with more concise labels for plotting
coef_all$term <- recode(coef_all$term,
  activity150_z = "Aerobic activity",
  strength_z = "Muscle strengthening",
  inactive_z = "No leisure activity",
  fruit_z = "Low fruit consumption",
  veg_z = "Low vegetable consumption"
)

# Plot posterior estimates and 95% credible intervals across models
ggplot(coef_all, aes(x = estimate, y = term)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~model, scales = "free_y") +
  labs(
    title = "Posterior Estimates of Behavioral Predictors on Obesity",
    x = "Estimated Effect on Obesity Prevalence (95% Credible Interval)",
    y = NULL
  ) +
  theme_minimal()
```

The figure summarizes posterior estimates and 95% credible intervals for behavioral predictors across, nutrition, activity, and inactivity. The points represent estimated posterior means of each predictor, horizontal lines represent the 95% credible interval, and the dashed vertical lines at 0 represent no association with obesity prevalence. 

Overall, the results show that aerobic activity is negatively associated with obesity prevalence, while lower fruit consumption and no physical activity during leisure time are positively associated with obesity prevalence. In contrast, predictors such as muscle strengthening activity and low vegetable consumption contain 0 in their credible intervals, which suggest weaker or more uncertain associations with obesity prevalence. 

## Conclusion

**What behavioral factors are posteriorly associated with obesity prevalence, including physical activity, sedentary behavior, and nutrition?**

The posterior results suggest that several behavioral factors are associated with obesity prevalence. Higher rates of leisure time inactivity and lower fruit consumption were positively associated with obesity prevalence. On the other hand, higher rates of aerobic activity were negatively associated with obesity prevalence. 

Factors such as muscle strengthening and vegetable consumption were more uncertain, providing limited evidence limited evidence for a clear association between obesity prevalence. 